Deploying and Sharing QSAR Models using Predictive Modeling Markup Language

### 2014-12-03 paper.tex
Model types in QSAR

>>>>

\subsection{Fitness of PMML for QSAR purposes}
Four major types of models:
\begin{itemize}
  \item Regression. Predicts continuous dependent value.
  Supervised method.
  \item Classification. Predicts categorical (unordered labels) or ordinal 
  (ordered labels) dependent values. Two categories (binomial response) or
  multiple categories (multinomial response). Supervised method.
  \item Clustering. Grouping instances by similarity. Unsupervised method.
  \item Association rules. Predicting the next instance in sequence.
  Supervised method.
\end{itemize}

Historically, research stated around regression applications in physical organic
chemistry such as predicting reaction rates and equilibrium constants (eg. Hammett 
equation) using linear relationships. Later expanded into more "fuzzy" 
application areas such as predicting biological activities (eg. Hansch equation)
where the relationship between the dependent variable and independent variables
is often non-linear or otherwise poorly understood (requires \textit{a posteriori} model 
interpretation). Classification models can be often regarded as simplifications
of regression models, where the range of a continuous response variable has been
discretized ("binned") into small number of categories. For example, for regulatory
purposes, it may be better to translate the predicted effect concentration based
on safety thresholds to "not hazardous", "moderately hazardous", "highly hazardous"
class labels. Of course, there are also many true classification problems, where
the goal is to assign the chemical compound to one predefined chemical category.
For example, classifying a toxicant by its structural features as "chemically inert",
"electrophile" or "nucleophile" (without regard to its effective concentration).
Such true classification models lead to clustering models, where the chemical space
is dissected to subregions based on similarity measures, either structural features
or molecular descriptor values. For example, dissecting a two-dimensional chemical
space into four quadrants. Clustering models are the basis for read-across methods,
where the characteristics of a new chemical compound are predicted from the 
characteristics of its closest analogues.
